---
title: "Data 607 - TidyVerse CREATE assignment"
author: "Shamecca Marshall"
date: "2023-10-15"
output:
  html_document: default
  pdf_document: default
---

In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository:  https://github.com/peterkowalchuk/FALL2023TIDYVERSE

FiveThirtyEight.com datasets.

Kaggle datasets. 

Your task here is to Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)

Later, you'll be asked to extend an existing vignette.  Using one of your classmate’s examples (as created above), you'll then extend his or her example with additional annotated code. (15 points)

## Load Libraries
```{r}
library(ggplot2)
library(dplyr)
library(ggthemes)
```

## Read CSV

I wanted to view the population, HIV Deaths, and life expectancy of Kenya throughout the years. Original data can be found at the link :
https://www.kaggle.com/datasets/amirhosseinmirzaie/countries-life-expectancy
```{r}
Life_expectancy <- read.csv("https://raw.githubusercontent.com/Meccamarshall/Data607/main/life_expectancy.csv")
colnames(Life_expectancy)
```
## Kenya Population
```{r}
Kenya_Population <- Life_expectancy%>%
filter(Country == "Kenya")%>%
ggplot(aes(Year, Population, group = 1)) + geom_point(na.rm=TRUE, color = "hotpink") + geom_line(na.rm=TRUE, color = "orange")+
  labs(title = "Kenya Population Throughout the Years", x = "Year", y = "Population (million)")
Kenya_Population
```

## Kenya HIV Deaths
```{r}
HIV_deaths_Kenya <- Life_expectancy%>%
filter(Country == "Kenya")%>%
ggplot(aes(Year, HIV.AIDS, group = 1)) + geom_point(na.rm=TRUE, color = "hotpink") + geom_line(na.rm=TRUE, color = "orange")+
  labs(title = "Kenya deaths caused by AIDS of the last 4-year-olds", x = "Year", y = "HIV/AIDS Deaths")
HIV_deaths_Kenya
```
## Kenya Life Expectancy 
```{r}
Kenya_Life_expectancy <- Life_expectancy%>%
filter(Country == "Kenya")%>%
ggplot(aes(Year, Life.expectancy, group = 1)) + geom_point(na.rm=TRUE, color = "hotpink") + geom_line(na.rm=TRUE, color = "orange")+
  labs(title = "Kenya Life Expenctancy Throughout the Years", x = "Year", y = "Life expectancy")
Kenya_Life_expectancy
```