---
title: "Tidyverse Assignment"
author: "Steve Phillips"
date: "2023-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(tidyverse)
```
## I will be showing a few dplyr functions that might be useful
```{r}
df <- read_csv("https://raw.githubusercontent.com/sphill12/DATA607/main/mlb_elo.csv")
```
```{r}
glimpse(df)
```
## Sometimes we want to apply transformations across multiple columns. The across function is best for manipulating multiple columns in conjunction with the mutate function. Here is an example, where we reduce the amount of decimal places that the data covers for elo ratings

```{r}
df %>% mutate(across(c(elo1_pre,elo2_pre), round))
```
## For easier analysis, sometimes you will want to arrange data in a certain way. The arrange function is useful for these purposes. Here we can see the first data points for baseball elo. The data starts in 1871.
```{r}
df %>% arrange((date))
```
## Sometimes it is ideal to create dummy variables for easy analysis. The case_when function can be used for these purposes
```{r}
df %>% mutate(is_sea = case_when(team1 == "SEA" ~ "Yes"))
```
## Another function to help with dummy variables is the if_else function. This allows for splitting things into two groups
```{r}
df %>% mutate(is_nym = if_else(team1 == "NYM", "Yes", "No"))
```


# Matthew Roland's Extensions

##Using the "Between" Function

When using a long time-series dataset such as this, you may want to only look at a subset of values. For instance, perhaps you don't want to view/analyze player information from the 1800s due to a lack of relevance, or the bearth of missing data unique to that time-period. To work around this, we can use dplyr's "between" function to subset a range of dates that we are interested in. Let's say we want to only examine the 2022 and 2023 seasons:

```{r}
str(df)

df_filt <- df %>% filter(between(season, 2022, 2023))

df_filt <- df_filt %>% arrange(date)
glimpse(df_filt)
```
As we can see, we are now left with a dataframe containing only observations from 2022 through 2023

## Pull Function

Maybe you would like to quickly pull up comparisons for each pitchers' scores. Fortunately, you can do this with dplyr's **pull** function, which serves as a neater substitute for native R's **$** operator: 

```{r}
#First, I'll find the average score for each 1st pitcher
score <- df_filt %>% group_by(pitcher1) %>% summarize(mean_score = mean(score1))

##And now, let's pull a list of each pitcher and their average scores
score %>% pull(mean_score, pitcher1)
```

Now we can quickly see the average score of each player