---
title: "Tidyverse Create Asignment"
author: "Noori Selina"
date: "2023-11-07"
output: pdf_document
---

```{r}
# Load necessary libraries
library(tidyverse)
```

# Introduction 

This vignette demonstrates how to perform data manipulation and analysis using the dplyr package. We will work with the Census Income dataset, which contains information on individuals' demographics and income levels.The data set can be found here: https://www.kaggle.com/datasets/tawfikelmetwally/census-income-dataset/data


# Data Loading 
To load the data, I will be importing the data from a GitHub link.
```{r}
census_data <- read_csv("https://raw.githubusercontent.com/NooriSelina/Data-607/main/censusincome.csv")
head(census_data)
```
# Data Manipulation with dplyr
The dplyr package provides a set of functions for data manipulation. We will use the following functions to explore the dataset. 

1. Filtering - Using the dplyr package, I will use the `filter()` function to filter the dataset based on specific criteria. In this case, we are interested in individuals with incomes greater than $50,000, so we filter the dataset to include only such individuals.

```{r}
high_income_data <- census_data %>%
  filter(Income == ">50K")
head(high_income_data)
```


2. Grouping and Summarizing - The group_by() and summarize() functions of the dplyr package are valuable for aggregating data. We will group the data by education level and calculate summary statistics for age and hours worked per week within each education category.

```{r}
income_summary <- high_income_data %>%
  group_by(Income) %>%
  summarize(
    mean_age = mean(Age, na.rm = TRUE),
    median_hours = median(`Hours per Week`, na.rm = TRUE)
  )

print(income_summary)
```

3. Sorting - The arrange() function is used to sort the summarized data. In this example, we sort the education summary by mean age in descending order, which helps identify the education categories with the highest mean age.

```{r}
high_income_data <- high_income_data %>%
  arrange(desc(Age))

print(head(high_income_data, 10))
```


# Conclusion
In this vignette, the `dplyr` package from the `tidyverse` collection was used to efficiently manage and analyze the Census Income dataset. Specifically, the data was filtered using the `filter()` function to focus on high-income individuals, grouped by income levels using the `group_by()` function, and then sorted by mean age with the `arrange()` function.

This approach uncovered valuable insights about the demographics of high-income individuals. The `dplyr` package, along with other tools from the `tidyverse`, made data manipulation tasks straightforward and effective.


### Haig Bedros' Changes

To extend Selina Noori's code I will introduce a few more `dplyr` functions that can be used for further data manipulation and analysis. 

Given the structure of your `census_data` dataset, it seems that there is no direct column to join with the `education_data` on a state basis. Since `census_data` does not have a `STATE` or similar column, we'll need to adapt our approach. One possibility is to use the `Native Country` field, but this doesn't seem to align well with the state-based `education_data`.

However, we can still demonstrate advanced `dplyr` techniques using just the `census_data`. Here's how the vignette can be adjusted:

---

```{r}
# Load the dplyr library, part of tidyverse
library(tidyverse)
```

## Advanced Data Manipulation with dplyr

### Mutating Data

Let's start by categorizing individuals based on their age. We can create a new column `age_group` to facilitate this analysis.

```{r}
census_data <- census_data %>%
  mutate(age_group = case_when(
    Age <= 25 ~ "Young",
    Age > 25 & Age <= 50 ~ "Adult",
    TRUE ~ "Senior"
  ))
```

### Summarizing Across Multiple Groups

Next, we can group the data by `age_group` and `Education` to understand the distribution of education levels across different age groups.

```{r}
education_age_summary <- census_data %>%
  group_by(age_group, Education) %>%
  summarize(
    average_hours = mean(`Hours per Week`, na.rm = TRUE),
    count = n(),
    median_income = median(`Final Weight`, na.rm = TRUE)
  )

print(education_age_summary)
```

### Exploring Gender and Workclass

We can also explore the relationship between gender, workclass, and income. This will help us understand how these factors interact.

```{r}
gender_workclass_summary <- census_data %>%
  group_by(Gender, Workclass) %>%
  summarize(
    high_income_count = sum(Income == ">50K"),
    total_count = n()
  )

print(gender_workclass_summary)
```

## Conclusion

In this revised vignette, we demonstrated the power of `dplyr` in manipulating and analyzing the `census_data` dataset. We created new categories, summarized data across multiple groups, and explored the relationships between gender, workclass, and income. These techniques from the `tidyverse` collection are crucial for effective data analysis and provide a foundation for more complex analytical tasks.
